Journey: A Study of Studying Large Memory Systems

### 2017-04-12 paper.tex
Added swapping experiments

@@

Our methodology is broken into two parts. First, we measure the scalability of
several Linux memory management system calls on a system running natively on
bare metal. Then, we measure the scalability of Linux virtual machines running
on the KVM hypervisor.

\subsection{Native Experiments}

@@

We measure the memory overhead and latency of the Linux Memory Management
system's use of \texttt{vm\_area\_struct}, which represents a region of a
process's virtual address space. (TODO: what about page tables? struct page?)
When a program makes a memory management system calls such as \texttt{mmap},
\texttt{mremap}, \texttt{munmap}, \texttt{mprotect}, \texttt{mlock}, or
\texttt{madvise}, the kernel is creating, changing, or removing
\texttt{vm\_area\_struct}s; thus, measuring the scalability of Linux memory
management requires measuring the overhead of these structs.

@@

\item \textit{Random}. This workload allocates $2^{20}$ randomly placed,
non-overlaping memory regions.

@@

\subsection{VM Experiments}

>>>>

(TODO: we might want to add stuff about using swap instead of larger memory)

@@

Our methodology is broken into two parts. First, we measure the scalability of
Linux's \texttt{vm\_area\_struct}. Then, we measure the minimum amount of memory
necessary for swapping.

\subsection{\texttt{vm\_area\_struct} Experiments}

@@

We measure the memory overhead and latency of the Linux Memory Management
system's use of \texttt{vm\_area\_struct}, which represents a region of a
process's virtual address space.  We do not measure the overhead of
\texttt{struct page} because it is known that there is one struct for each
physical page. We do not measure the overhead of page tables because this can be
calculated from the size of the virtual memory space used by the process.  When
a program makes a memory management system calls such as \texttt{mmap},
\texttt{mremap}, \texttt{munmap}, \texttt{mprotect}, \texttt{mlock}, or
\texttt{madvise}, the kernel is creating, changing, or removing
\texttt{vm\_area\_struct}s; thus, measuring the scalability of Linux memory
management requires measuring the overhead of these structs.

@@

\item \textit{Random}. This workload allocates randomly the pages in a
inside of a $2^{20}$ page region.

@@

\subsection{Swapping Experiments}

These experiments are designed to measure the minimum amount of memory needed
to using swapping.

In this experiment, a single process touches 32GB of virtual memory, leading to
the actual allocation of physical memory. Notice that 32GB is twice as large as
the 16GB of physical memory the test machine. This causes Linux kernel to start
swapping out pages to make more room.

After each page of virtual memory is touched, the resident set size (RSS; the
amount of physical memory Linux has allocated to the process) is recorded. This
value is graphed in our results section.

We repeat this experiment with 5 different swapfile sizes to capture the impact
of swapfile size on swapping overhead.

(TODO: strided access vs random access?)

\section{Result}

\subsection{\texttt{vm\_area\_struct} Experiments}

\subsection{Swapping Experiments}
